---
title: "Chapter 12 Problem 2"
author: "Bri Bentley"
date: '`r format(Sys.time(), "%b %d, %Y")`'
output:
  bookdown::html_document2:
    highlight: textmate
    theme: yeti
---

**A metallurgic company is investigating lost revenue due to worker illness. It is interested in creating a table of lost revenue to be used for future budgets and company forecasting plans. The data are stored in the data frame** `LOSTR`.

```{r}
library(PASWR2)
summary(LOSTR)
head(LOSTR)
```


a) Create a scatterplot of lost revenue versus number of ill workers. Characterize the relationship between `lostrevenue` and `numbersick`.

There is a postive correlation between `lostrevenue` and `numbersick`. There also appears that there may be a single outlier.

b) Fit a regression line to the data. Superimpose the resulting line in the plot created in part (a).

```{r}
library(ggplot2)
ggplot(data = LOSTR, aes(y = lostrevenue, x = numbersick)) +
  geom_point() + 
  geom_smooth(method = "lm")
```

c) Compute the covariance matrix of the B-hats.

```{r}
mod <- lm(lostrevenue ~ numbersick, data = LOSTR)
X <- model.matrix(mod)
dim(X)

anova(mod)
SSE <- anova(mod)[2,2]
MSE <- SSE/(25-2)

MSE * solve(t(X) %*% X)
```

d) Create a 95% confidence interval for B1.

```{r}
confint(mod, level = 0.95)
```

e) Compute the coefficient of determination and the adjusted coefficient of determination. Provide contextualual interpretations of both values.

```{r}
summary(mod)$r.squared
summary(mod)$adj.r.squared
```

The coefficient of determination is 0.8845 and the adjusted coefficient of determination is 0.8795. 

About 88 percent of the variation in the association between number of sick employees and the amount of lost revenue has been explained by the model.

f) What assumptions need to be satisfied in order to use the model from part (b) for inferential purposes? If there is/are any outlier/s in the data, remove it/them prior to answering the remainder of the questions.

We assume that their is no influential data point that affects our regression model.

There is an outlier, data point 4: (11, 2500)

```{r}
library(car)
outlierTest(mod)
```

```{r}
#Removing the outlier
library(dplyr)
library(tidyverse)
newdata <- LOSTR %>% slice(-4)
newdata
```

g) Determine the expected lost revenues when 5, 15, and 25 workers are absent due to illness.

```{r}
mod2 <- lm(lostrevenue ~ numbersick, data = newdata)
summary(mod2)
ycep <- summary(mod2)$coefficients[1,1]
slope <- summary(mod2)$coefficients[2,1]
ycep
slope

ExpLostRev <- slope * c(5, 15, 25) + ycep
ExpLostRev
```

The expected lost revenues when 5, 15, and 25 workers are absent due to illness is `r ExpLostRev `.

h) Compute a 95% prediction interval of lost revenues when 14 workers are absent due to illness.
```{r}
confint(mod2, level = 0.95)
```

