---
title: 'Quiz #2 - Partial Solution'
author: "Your Name"
date: 'Last knit on `r format(Sys.time(), "%B %d, %Y at %X")`'
output: 
  bookdown::html_document2
---

```{r label = "setup", include=FALSE}
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.align = "center", comment = NA, fig.height = 3.5)
```

**Directions: Change the author field in the `YAML` from "Your Name" to your name using the convention of FirstName LastName. For example, my entry would be `author: "Alan Arnholt"`.  The data frame `ncbirths` contains the variables `fage`, `mage`, `mature`, `weeks`, `premie`, `visits`, `marital`, `gained`, `weight`, `lowbirthweight`, `gender`, `habit`, and `whitemom`.  Store this file as `Quiz02.Rmd` inside your `Quizzes` directory.  When you finish, commit your changes and push your work to GitHub.  Use the commit message `Finished Quiz 02!`.  Verify you can see your files at `https://github.com/STAT-ATA-ASU/YourLast-YourFirst-STT5812-SP18/tree/master/Quizzes`.**

```{r}
library(openintro)
names(ncbirths)
# Type ?ncbirths 
# in the console to open the help file
# I like to use summary() and str() on all new data sets
summary(ncbirths)
str(ncbirths)
```

1. Compute mean, standard deviation, and number of non-missing observations for the weight gained during pregnancy (`gained`) by mothers according to their level of `mature` (`mature mom` or `younger mom`).

```{r}
library(dplyr)
SI <- ncbirths %>% 
  group_by(mature) %>% 
  summarize(Mean = mean(gained, na.rm = TRUE), SD = sd(gained, na.rm = TRUE), n = sum(!is.na(gained)))
SI
## Using tapply - will have to use an annonymous function
tapply(ncbirths$gained, ncbirths$mature, mean, na.rm = TRUE)
with(ncbirths, tapply(gained, mature, sd, na.rm = TRUE))
tapply(ncbirths$gained, ncbirths$mature, function(x) {length(x[!is.na(x)])})
tapply(ncbirths$gained, ncbirths$mature, function(x) {sum(!is.na(x))})
## Using subset---could use drop = TRUE to make vector
MM <- subset(ncbirths, subset = mature == "mature mom", select = gained)
YM <- subset(ncbirths, subset = mature == "younger mom", select = gained)
MX <- mean(MM$gained, na.rm = TRUE)
SX <- sd(MM$gained, na.rm = TRUE)
nx <- sum(!is.na(MM$gained))
MY <- mean(YM$gained, na.rm = TRUE)
SY <- sd(YM$gained, na.rm = TRUE)
ny <- sum(!is.na(YM$gained))
c(MX, SX, nx, MY, SY, ny)
## Using pure R
MMg <- ncbirths$gained[ncbirths$mature == "mature mom"]
YMg <- ncbirths$gained[ncbirths$mature == "younger mom"]
mean(MMg, na.rm = TRUE)
sd(MMg, na.rm = TRUE)
sum(!is.na(MMg))
mean(YMg, na.rm = TRUE)
sd(YMg, na.rm = TRUE)
sum(!is.na(YMg))
```

The mean, standard deviation, and number of non-missing observations for mature mothers and younger mothers are: $`r SI[1,2]`, `r SI[1,3]`, `r SI[1,4]`, `r SI[2,2]`, `r SI[2,3]`, `r SI[2,4]`$, respectively.

2. Based on your answers to question 1, the density plots shown in Figure \@ref(fig:dens), and the Q-Q plots shown in Figure \@ref(fig:qqp) 
    + Should you pool or not pool the variances to compute a confidence interval for the mean weight gained during pregnancy for mature moms minus the mean weight gained during pregnancy for younger mothers  $\left( \mu_{\text{mature mom}} - \mu_{\text{younger mom}} \right)$?  
    + Construct a 90% confidence interval for the mean weight gained during pregnancy for mature moms minus the mean weight gained during pregnancy for younger mothers  $\left( \mu_{\text{mature mom}} - \mu_{\text{younger mom}} \right)$.  
    + Does your confidence interval contain zero?  If so, what does this mean? 


```{r, label = "dens", fig.cap = "Density plots of weight gained (lbs.)", echo = FALSE}
library(ggplot2)
ggplot(data = ncbirths, aes(x = gained, fill = mature)) +
  geom_density() + 
  theme_bw() + 
  facet_grid(.~mature) + 
  guides(fill = FALSE)
```

```{r, label = "qqp", fig.cap = "Q-Q plot of weight gained by `mature`", echo = FALSE}
ggplot(data = ncbirths, aes(sample = gained, color = mature)) +
  geom_qq(size = 1, alpha = 0.2) + 
  facet_grid(.~mature) + 
  theme_bw() + 
  guides(color = FALSE)
```

```{r}
# Your answer here
CI <- t.test(gained ~ mature, data = ncbirths, conf.level = 0.9)$conf
CI
# Could also pass two separate vectors
t.test(MMg, YMg, conf.level = 0.9)$conf
# Not recommended but since you have all the information
# one could use tsum.test
PASWR2::tsum.test(mean.x = MX, mean.y = MY, s.x = SX, s.y = SY, n.x = nx, n.y = ny, conf.level = 0.9)$conf
```

* Your Answers:
    + **Do not** pool the variances since the larger standard deviation is associated with the larger sample size and the sample sizes are different.  One might argue to pool the variances since a confidence interval for the ratio of variances contains 1; however, this will produce more errors than not pooling.
    + A 90% confidence interval for the mean weight gained during pregnancy for mature moms minus the mean weight during preganancy gained for younger mothers  $\left( \mu_{\text{mature mom}} - \mu_{\text{younger mom}} \right)$ is $CI_{0.90}\left( \mu_{\text{mature mom}} - \mu_{\text{younger mom}} \right) = [`r CI[1]`\text{ pounds}, `r CI[2]`\text{ pounds}]$.
    + Since the confidence interval **contains** zero, this suggests that there is **no significant** difference between the mean weight gained during pregnancy for mature mothers and the mean weight gained during pregnancy for younger mothers.

3. Construct a 95% Wilson confidence interval for the proportion of mature mothers in North Carolina.

```{r}
xtabs(~mature, data = ncbirths)
# Your code here
library(binom)
CI <- binom.confint(x = 133, n = 133 + 867, conf.level = 0.95, methods = "wilson")
CI
# Or
prop.test(x = 133, n = 133 + 867, correct = FALSE)
```

The 95% Wilson confidence interval for the proportion of mature mothers in North Carolina is $CI_{0.95}(\pi) = [`r CI[5]`, `r CI[6]`]$.


4.  Compute a 90% confidence interval for the variance in weight gained during pregnancy by North Carolina mature mothers.


```{r}
# Your code here
# 90% CI for sigma^2
CI <- ncbirths %>% 
  filter(mature == "mature mom") %>% 
  summarize(Mean = mean(gained, na.rm = TRUE), SD = sd(gained, na.rm = TRUE), n = sum(!is.na(gained))) %>% 
  mutate(lcl = (n - 1)*SD^2/qchisq(0.95, n - 1), ucl = (n - 1)*SD^2/qchisq(0.05, n - 1))
CI
# Or
ACI <- c((nx - 1)*SX^2/qchisq(0.95, nx - 1), (nx -1)*SX^2/qchisq(0.05, nx - 1))
ACI
```

The 90% confidence interval for the variance in weight gained during pregnancy by North Carolina mature mothers is $CI_{0.90}(\sigma^2) = [`r CI[4]` \text{ pounds}^2, `r CI[5]` \text{ pounds}^2]$.