---
title: 'Quiz #1'
author: "Bri Bentley"
date: 'Last knit on `r format(Sys.time(), "%B %d, %Y at %X")`'
output: html_document
---

```{r label = "setup", include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.align = "center", comment = NA)
```

**Directions: Change the author field in the `YAML` from "Your Name" to your name using the convention of FirstName LastName. For example, my entry would be `author: "Alan Arnholt"`.  The data frame `Y1980` contains the variables `teamID`, `W`, `L`, and `attendance` for the year 1980.  Create a `Quizzes` directory inside your private repository.  Store this file as `Quiz01.Rmd` inside the `Quizzes` directory.  When you finish, commit your changes and push your work to GitHub.  Use the commit message `Finished Quiz 01!`.  Verify you can see your files at `https://github.com/STAT-ATA-ASU/YourLast-YourFirst-STT5812-SP18/tree/master/Quizzes`.**

```{r}
library(Lahman)
library(dplyr)
Y1980 <- Teams %>%
  filter(yearID == 1980) %>% 
  select(teamID, W, L, attendance)
head(Y1980)  # show frist six rows
```

1. Create a histogram or density plot showing the distribution of the variable `attendance` from the data frame `Y1980`.  

```{r}
# Your Code here
hist(Y1980$attendance, breaks = 8, main = "Histogram of Attendance", xlab = "Attendance")
```

2. Create a quantile-quantile plot of the variable `attendance` from the data frame `Y1980`.

```{r}
# Your code here
qqnorm(Y1980$attendance)
qqline(Y1980$attendance)
```

3.  Assume the conditions for constructing a confidence interval for $\mu$ based on a $t$ distribution are satisfied.  Find a report the following values assuming $\alpha = 0.10$:

```{r, echo = FALSE, results = "hide"}
# You will need code here if you want to answer using inline R
t.test(Y1980$attendance, data = Y1980, conf.level = .90)
CI <- t.test(Y1980$attendance, data = Y1980, conf.level = .90)$conf
s <- sd(Y1980$attendance)
s
n <- length(Y1980$attendance)
s
xbar <- mean(Y1980$attendance)
```


  a. $\bar{x} = `r xbar`$
  b. $s = `r s`$
  c. $n = `r n`$
  d. $t_{1 - \alpha/2,\, n -1} = 12.889$
  e. $CI_{1 - \alpha}(\mu) = (`r CI[1]`, `r CI[2]`)$
 
**Extra Credit: if you use inline R to answer a-e.**

## Instructor Comments    

* No spaces in document names.

* 3 d. requests $t_{0.95, 25} = `r qt(0.95, 25)`$

* Inline code is nice!