---
title: 'Quiz #1 - Partial Solution'
author: "Your Name"
date: 'Last knit on `r format(Sys.time(), "%B %d, %Y at %X")`'
output: html_document
---

```{r label = "setup", include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.align = "center", comment = NA)
```

**Directions: Change the author field in the `YAML` from "Your Name" to your name using the convention of FirstName LastName. For example, my entry would be `author: "Alan Arnholt"`.  The data frame `Y1980` contains the variables `teamID`, `W`, `L`, and `attendance` for the year 1980.  Create a `Quizzes` directory inside your private repository.  Store this file as `Quiz01.Rmd` inside the `Quizzes` directory.  When you finish, commit your changes and push your work to GitHub.  Use the commit message `Finished Quiz 01!`.  Verify you can see your files at `https://github.com/STAT-ATA-ASU/YourLast-YourFirst-STT5812-SP18/tree/master/Quizzes`.**

```{r}
library(Lahman)
library(dplyr)
Y1980 <- Teams %>%
  filter(yearID == 1980) %>% 
  select(teamID, W, L, attendance)
head(Y1980)  # show frist six rows
```

1. Create a histogram or density plot showing the distribution of the variable `attendance` from the data frame `Y1980`.  

```{r}
# Your Code here
library(ggplot2)
ggplot(data = Y1980, aes(x = attendance)) + 
  geom_density(fill = "purple", alpha = 0.3) +
  theme_bw()
```

2. Create a quantile-quantile plot of the variable `attendance` from the data frame `Y1980`.

```{r}
# Your code here
ggplot(data = Y1980, aes(sample = attendance)) +
  stat_qq() + 
  theme_bw()
shapiro.test(Y1980$attendance)
```

3.  Assume the conditions for constructing a confidence interval for $\mu$ based on a $t$ distribution are satisfied.  Find a report the following values assuming $\alpha = 0.10$:

```{r, echo = FALSE, results = "hide"}
# My code....remove
alpha <- 0.10
xbar <- mean(Y1980$attendance)
s <- sd(Y1980$attendance)
n <- sum(!is.na(Y1980$attendance))
CT <- qt(1 - alpha/2, n - 1)
SE <- s/sqrt(n)
ME <- CT*SE
lcl <- xbar - ME
ucl <- xbar + ME
round(c(xbar, s, n, CT, SE, ME, lcl, ucl), 3)
```


  a. $\bar{x} =`r round(xbar, 3)`$
  b. $s =`r round(s, 3)`$
  c. $n =`r n`$
  d. $t_{1 - \alpha/2,\, n -1} =`r round(CT, 3)`$
  e. $CI_{1 - \alpha}(\mu) =[`r round(lcl, 3)`, `r round(ucl, 3)`]$
 
**Extra Credit: if you use inline R to answer a-e.**

```{r}
# Your code here
alpha <- 0.10
xbar <- mean(Y1980$attendance)
s <- sd(Y1980$attendance)
n <- sum(!is.na(Y1980$attendance))
CT <- qt(1 - alpha/2, n - 1)
SE <- s/sqrt(n)
ME <- CT*SE
lcl <- xbar - ME
ucl <- xbar + ME
round(c(xbar, s, n, CT, SE, ME, lcl, ucl), 4)
t.test(Y1980$attendance, conf.level = 0.90)$conf
# Using tidyverse
Y1980 %>% 
  summarize(xbar = mean(attendance), s = sd(attendance), n = sum(!is.na(attendance))) %>% 
  mutate(CT = qt(1 - alpha/2, n - 1), lcl = xbar - CT*s/sqrt(n), ucl = xbar +CT*s/sqrt(n))
```
    
