---
title: "Chapter 12 Problem 2"
author: "Bri Bentley"
date: '`r format(Sys.time(), "%b %d, %Y")`'
output:
  bookdown::html_document2:
    highlight: textmate
    theme: yeti
---

**A metallurgic company is investigating lost revenue due to worker illness. It is interested in creating a table of lost revenue to be used for future budgets and company forecasting plans. The data are stored in the data frame** `LOSTR`.

```{r}
library(PASWR2)
summary(LOSTR)
head(LOSTR)
```


a) Create a scatterplot of lost revenue versus number of ill workers. Characterize the relationship between `lostrevenue` and `numbersick`.

There is a postive correlation between `lostrevenue` and `numbersick`. There also appears that there may be a single outlier.

b) Fit a regression line to the data. Superimpose the resulting line in the plot created in part (a).

```{r}
library(ggplot2)
ggplot(data = LOSTR, aes(y = lostrevenue, x = numbersick)) +
  geom_point() + 
  geom_smooth(method = "lm")
```

c) Compute the covariance matrix of the B-hats.

```{r}
mod <- lm(lostrevenue ~ numbersick, data = LOSTR)
X <- model.matrix(mod)
dim(X)

vcov(mod)
```

d) Create a 95% confidence interval for B1.

```{r}
CI <- confint(mod, level = 0.95)
CI
```

The confidence interval is $[`r CI[2,1]`, `r CI[2,2]`]$.

e) Compute the coefficient of determination and the adjusted coefficient of determination. Provide contextualual interpretations of both values.

```{r}
summary(mod)$r.squared
summary(mod)$adj.r.squared
```

The coefficient of determination is 0.8845 and the adjusted coefficient of determination is 0.8795. 

The coefficient of determination and adjusted coefficient of determination means that about 88.5% and 88%, respectively, of the variance in revenue lost can be explained by the change in `numbersick`.

f) What assumptions need to be satisfied in order to use the model from part (b) for inferential purposes? If there is/are any outlier/s in the data, remove it/them prior to answering the remainder of the questions.

We assume that the errors are assumed to be independent, follow a normal distribution with a mean of zero, and have constant variance.

There is an outlier, data point 4: $(11, 2500)$

```{r}
library(car)
outlierTest(mod)
```

```{r warning = FALSE, message = FALSE}
#Removing the outlier
library(dplyr)
library(tidyverse)
newdata <- LOSTR %>% slice(-4)
newdata
```

g) Determine the expected lost revenues when 5, 15, and 25 workers are absent due to illness.

```{r}
mod2 <- lm(lostrevenue ~ numbersick, data = newdata)
summary(mod2)
ycep <- summary(mod2)$coefficients[1,1]
slope <- summary(mod2)$coefficients[2,1]
ycep
slope

ExpLostRev <- slope * c(5, 15, 25) + ycep
ExpLostRev
```

The expected lost revenues when 5, 15, and 25 workers are absent due to illness is `r ExpLostRev ` respectively.

h) Compute a 95% prediction interval of lost revenues when 14 workers are absent due to illness.
```{r}
PI <- predict(mod2, data.frame(numbersick = 14), interval = "predict")
PI
```

The 95% prediction interval of lost revenue when 14 workers are absent due to illness is between $(`r PI[1,2]`, `r PI[1,3]`)$.