---
title: 'Quiz #2'
author: "Bri Bentley"
date: 'Last knit on `r format(Sys.time(), "%B %d, %Y at %X")`'
output: 
  bookdown::html_document2
---

```{r label = "setup", include=FALSE}
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.align = "center", comment = NA, fig.height = 3.5)
```

**Directions: Change the author field in the `YAML` from "Your Name" to your name using the convention of FirstName LastName. For example, my entry would be `author: "Alan Arnholt"`.  The data frame `ncbirths` contains the variables `fage`, `mage`, `mature`, `weeks`, `premie`, `visits`, `marital`, `gained`, `weight`, `lowbirthweight`, `gender`, `habit`, and `whitemom`.  Store this file as `Quiz02.Rmd` inside your `Quizzes` directory.  When you finish, commit your changes and push your work to GitHub.  Use the commit message `Finished Quiz 02!`.  Verify you can see your files at `https://github.com/STAT-ATA-ASU/YourLast-YourFirst-STT5812-SP18/tree/master/Quizzes`.**

```{r}
library(openintro)
names(ncbirths)
# Type ?ncbirths 
# in the console to open the help file
```

```{r}
summary(ncbirths)
```

1. Compute mean, standard deviation, and number of non-missing observations for the weight gained during pregnancy (`gained`) by mothers according to their level of `mature` (`mature mom` or `younger mom`).

```{r}
library(dplyr)
ncbirths %>%
  group_by(mature) %>%
  summarize(Means = mean(gained, na.rm = TRUE), SD = sd(gained, na.rm = TRUE), n = sum(!is.na(gained)))
```

The mean, standard deviation, and number of non-missing observations for mature mothers and younger mothers are: **28.79, 13.48, 129, 30.56, 14.35, 844**, respectively.

2. Based on your answers to question 1, the density plots shown in Figure \@ref(fig:dens), and the Q-Q plots shown in Figure \@ref(fig:qqp) 
    + Should you pool or not pool the variances to compute a confidence interval for the mean weight gained during pregnancy for mature moms minus the mean weight gained during pregnancy for younger mothers  $\left( \mu_{\text{mature mom}} - \mu_{\text{younger mom}} \right)$?  
    + Construct a 90% confidence interval for the mean weight gained during pregnancy for mature moms minus the mean weight gained during pregnancy for younger mothers  $\left( \mu_{\text{mature mom}} - \mu_{\text{younger mom}} \right)$.  
    + Does your confidence interval contain zero?  If so, what does this mean? 


```{r, label = "dens", fig.cap = "Density plots of weight gained (lbs.)", echo = FALSE}
library(ggplot2)
ggplot(data = ncbirths, aes(x = gained, fill = mature)) +
  geom_density() + 
  theme_bw() + 
  facet_grid(.~mature) + 
  guides(fill = FALSE)
```


```{r, label = "qqp", fig.cap = "Q-Q plot of weight gained by `mature`", echo = FALSE}
ggplot(data = ncbirths, aes(sample = gained, color = mature)) +
  geom_qq(size = 1, alpha = 0.2) + 
  facet_grid(.~mature) + 
  theme_bw() + 
  guides(color = FALSE)
```


```{r}
CI <- t.test(gained ~ mature, data = ncbirths, conf.level = 0.90)$conf
CI
```

* Your Answers:
    + **Do not** pool the variances.
    + A 90% confidence interval for the mean weight gained during pregnancy for mature moms minus the mean weight during preganancy gained for younger mothers  $\left( \mu_{\text{mature mom}} - \mu_{\text{younger mom}} \right)$ is $CI_{0.90}\left( \mu_{\text{mature mom}} - \mu_{\text{younger mom}} \right) = [`r CI[1]`,`r CI[2]`]$.
    + Since the confidence interval **contains** zero, this suggests that there **is no significant** difference between the mean weight gained during pregnancy for mature mothers and the mean weight gained during pregnancy for younger mothers.

3. Construct a 95% Wilson confidence interval for the proportion of mature mothers in North Carolina.

```{r}
library(binom)
xtabs(~mature, data = ncbirths)
binom.confint(x = 133, n = (133 + 867), methods = "wilson")
```

The 95% Wilson confidence interval for the proportion of mature mothers in North Carolina is $CI_{0.95}(\pi) = [0.1133512, 0.1554577]$.


4.  Compute a 90% confidence interval for the variance in weight gained during pregnancy by North Carolina mature mothers.


```{r}
# Your code here
# 90% CI for sigma^2
#Incorrect Code
#LCL <- ((133+867-1)*(13.48244)^2)/(qchisq(.05,133+867-1))
#UCL <- ((133+867-1)*(13.48244)^2)/(qchisq(.95,133+867-1))
#LCL
#UCL

#Corrected Code
C.I. <- ncbirths %>%
  filter(mature == "mature mom") %>%
  summarize(Mean = mean(gained, na.rm = TRUE), SD = sd(gained, na.rm = TRUE), n=sum(!is.na(gained))) %>%
  mutate(lcl = (n-1)*SD^2/qchisq(0.95, n-1), ucl = (n-1)*SD^2/qchisq(0.05, n-1))
C.I.
```

The 90% confidence interval for the variance in weight gained during pregnancy by North Carolina mature mothers is $CI_{0.90}(\sigma^2) = [149.721, 226.1878]$.

-------

In the first question I was not sure what to add to the n part of the code in order for it to discard the NA values in the total counts. I added that argument into my code in my revised version of the quiz. I also thought that since the graphs of the mature moms versus younger moms were so similar then it was appropriate to pool the variances, I was mistaken. The explanation you provided makes a lot of sense in terms of why my thinking on that problem was incorrect.

For the last problem, I made several errors. I forgot that the problem was specifically for mature moms, so when I found n I was basing it off of the entire population including the NA entries, and not specifically the mature mom population without the NA entries. I also used 0.05 for the lower confidence level and 0.95 for the upper confidence level when I should have switched those.