---
title: 'Quiz #3'
author: "Bri Bentley"
date: 'Last knit on `r format(Sys.time(), "%B %d, %Y at %X")`'
output: 
  bookdown::html_document2
---

```{r label = "setup", include=FALSE}
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.align = "center", comment = NA, fig.height = 3.5)
```

**Directions: Change the author field in the `YAML` from "Your Name" to your name using the convention of FirstName LastName. For example, my entry would be `author: "Alan Arnholt"`.   Store this file as `Quiz03.Rmd` inside your `Quizzes` directory.  When you finish, commit your changes and push your work to GitHub.  Use the commit message `Finished Quiz 03!`.  Verify you can see your files at `https://github.com/STAT-ATA-ASU/YourLast-YourFirst-STT5812-SP18/tree/master/Quizzes`.**

* Using a 5% significance level, what is the power of the test $H_0: \mu = 100$ versus $H_1: \mu > 100$ if a sample of size 36 is taken from a $N(105, 12)$?  

```{r}
z <- qnorm(.05, 100, 12/sqrt(36), lower.tail = FALSE)
beta <- pnorm(z, 105, 12/sqrt(36))
POWER <- 1 - beta
POWER
```
