---
title: 'Quiz #4'
author: "Bri Bentley"
date: 'Last knit on `r format(Sys.time(), "%B %d, %Y at %X")`'
output: 
  bookdown::html_document2
---

```{r label = "setup", include=FALSE}
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.align = "center", comment = NA, fig.height = 3.5)
```

**Directions: Change the author field in the `YAML` from "Your Name" to your name using the convention of FirstName LastName. For example, my entry would be `author: "Alan Arnholt"`.   Store this file as `Quiz04.Rmd` inside your `Quizzes` directory.  When you finish, commit your changes and push your work to GitHub.  Use the commit message `Finished Quiz 04!`.  Verify you can see your files at `https://github.com/STAT-ATA-ASU/YourLast-YourFirst-STT5812-SP18/tree/master/Quizzes`.**


A new drug is being developed to lower the total cholestrol (measured in milligrams per deciliter or mg/dL) for a group of high risk patients whose total cholestrol values follow a normal distribution with an average total cholestrol of 300 mg/dL and a standard deviation for total cholestrol of 50mg/dL. The researchers working on the drug want to detect a 25 mg/dL reduction in cholestrol at least 80% of the time while controling the type one error rate at 0.01.  You have been hired as the statistical consultant for the project.

1.  Specify the null and alternative hypotheses you will use to test the new drug.

    $H_0:$ $\mu$ = 300

    $H_1:$ $\mu$ < 300

2.  How many patients should you include in your sample? 
    a.  Use a *z-test* to answer the question. 317 (using shiny power app)
    b.  Use a *t-test* to answer the question. 43 (using shiny power app)
    
Check for power for the z distribution
```{r}
library(pwr)
pwr.norm.test(d = 25, n = NULL, sig.level = 0.01, power = 0.80, alternative = "less")
```

Check of the power for the t distribution.   
```{r}
std <- 50
delt <- 25
nt <- 43
power.t.test(nt, delt, std, sig.level = 0.01, type = "one.sample", alternative = "one.sided")
```

3.  Explain the assumption you made when using a *z-test* to answer question 2. that was not needed to use a *t-test*.  Do you think this assumption is reasonable? 

We assumed that our data was normal. I _____ think this assumption is reasonable given the graphs below.

```{r}
library(ggplot2)
ggplot(data = , aes()) +
  stat_qq() +
  theme_bw()
```

4.  Before the drug is administered, the lead physician states she believes the standard deviation for the patients administered the new drug will have a total cholestrol standard deviation between 40 mg/dL and 70 mg/dL.  Use this new information to recompute your sample size requirements for the experiment using a *t-test*.  Develop new recommendations and explain your new sample size requirements to meet the researchers' objectives.

    a.  Explain the costs associated with making type I and type II errors in this experiment.  Is it possible simultaneously to reduce both type I and type II errors?
    
    b.  What sample size is need to ensure the type II error is no greater than the type I error?

